---
title: "Formula one analysis"
output:
  html_document:
    self_contained: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    df_print: "default"
urlcolor: blue
---

```{r setup, include = FALSE}
# Global chunk options: hide code, suppress messages/warnings
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```
**Author**  
Leslie Huh, Nick Jitjang, Shria Parupudi, Sohum Walavalkar


***
***


# Project Overview

### Motivation
Detailed item (rated 0 to 4). What is the quality of these elements? Does the title tell us what to
expect of the report to follow? Does the motivation section convince the reader that the broader study
of this data set could elicit interesting insights?

### Dataset Description

Detailed item (rated 0 to 5). What is the quality of this description? Does it clearly communicate what
the rows and columns (i.e., observations and variables) are in the dataset in a way that is understandable
to a CMU statistics undergraduate audience?


### Research Questions

Detailed item (rated 0 to 5). What is the quality of the research questions? Are they well-motivated by
real-world/scientific interests? Or are they shallow? For example, a very shallow question would be
in the form, “what does the distribution of this variable look like?” - three questions like this would
guarantee 0pts for this item. A more interesting question would motivate why we would like to inspect
particular distributions with real-world context.

RQ1: [Question phrased in an interesting, real-world way, not just “what is the distribution of X?”]

RQ2: [Question]

RQ3: [Question]


***
***


# Research Question 1: 

### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Conclusion

[[conclustion for the research question]]

## Further Research

[[further reasrch for resarch question]]

# Research Question 2: 
### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Conclusion

[[conclustion for the research question]]

### Further Research

[[further reasrch for resarch question]]

# Research Question 3: 
### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Conclusion

[[conclustion for the research question]]

### Further Research

[[further reasrch for resarch question]]

# Research Question 4: 
### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Visualization 1

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]

### Conclusion

[[conclustion for the research question]]

### Further Research

[[further reasrch for resarch question]]

# Wrapping up





