---
title: "Formula one analysis"
output:
  html_document:
    self_contained: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    df_print: "default"
urlcolor: blue
---

```{r setup, include = FALSE}
# Global chunk options: hide code, suppress messages/warnings
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```
**Author**  
Leslie Huh, Nick Jitjang, Shria Parupudi, Sohum Walavalkar


***
***


# Project Overview

### Motivation
Detailed item (rated 0 to 4). What is the quality of these elements? Does the title tell us what to
expect of the report to follow? Does the motivation section convince the reader that the broader study
of this data set could elicit interesting insights?

### Dataset Description

Detailed item (rated 0 to 5). What is the quality of this description? Does it clearly communicate what
the rows and columns (i.e., observations and variables) are in the dataset in a way that is understandable
to a CMU statistics undergraduate audience?


### Research Questions

Detailed item (rated 0 to 5). What is the quality of the research questions? Are they well-motivated by
real-world/scientific interests? Or are they shallow? For example, a very shallow question would be
in the form, “what does the distribution of this variable look like?” - three questions like this would
guarantee 0pts for this item. A more interesting question would motivate why we would like to inspect
particular distributions with real-world context.

RQ1: [Question phrased in an interesting, real-world way, not just “what is the distribution of X?”]

RQ2: [Question]

RQ3: [Question]


***
***


# Research Question 1: 

### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1
Detailed item (0 to 5). What is the quality of these graphs? Are these graphs easily readable,
interpretable, and properly labeled?
• Detailed item (0 to 5). Are the graphs well-motivated, given the research questions of the project?
In other words, do the graphs address/answer the research questions of the project, or do they only
provide tangential (or even irrelevant) details?
[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
 Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?
### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?

### Conclusion

Does the report provide some clear conclusions that can be made from their
graphs and analyses?
• Detailed item (0 to 6). What is the accuracy of the claims being made? Are the claims well-supported
by the graphs and analyses presented? Are the conclusions well-aligned with the research questions of
the project?

### Further Research

Does the report discuss questions that have not been answered by the
project, but could be answered with future work?
• Detailed item (0 to 4). Did the team provide adequate reasons as to why these questions were left as
future work (e.g., they need more data, need more nuanced statistical techniques they haven’t learned,
etc.)? Are these future-work questions well-motivated given what the team has completed for this
project?

***
***

# Research Question 2: 

### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1
Detailed item (0 to 5). What is the quality of these graphs? Are these graphs easily readable,
interpretable, and properly labeled?
• Detailed item (0 to 5). Are the graphs well-motivated, given the research questions of the project?
In other words, do the graphs address/answer the research questions of the project, or do they only
provide tangential (or even irrelevant) details?
[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
 Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?
### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?

### Conclusion

Does the report provide some clear conclusions that can be made from their
graphs and analyses?
• Detailed item (0 to 6). What is the accuracy of the claims being made? Are the claims well-supported
by the graphs and analyses presented? Are the conclusions well-aligned with the research questions of
the project?

### Further Research

Does the report discuss questions that have not been answered by the
project, but could be answered with future work?
• Detailed item (0 to 4). Did the team provide adequate reasons as to why these questions were left as
future work (e.g., they need more data, need more nuanced statistical techniques they haven’t learned,
etc.)? Are these future-work questions well-motivated given what the team has completed for this
project?

***
***

# Research Question 3: 
### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1
Detailed item (0 to 5). What is the quality of these graphs? Are these graphs easily readable,
interpretable, and properly labeled?
• Detailed item (0 to 5). Are the graphs well-motivated, given the research questions of the project?
In other words, do the graphs address/answer the research questions of the project, or do they only
provide tangential (or even irrelevant) details?
[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
 Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?
### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?

### Conclusion

Does the report provide some clear conclusions that can be made from their
graphs and analyses?
• Detailed item (0 to 6). What is the accuracy of the claims being made? Are the claims well-supported
by the graphs and analyses presented? Are the conclusions well-aligned with the research questions of
the project?

### Further Research

Does the report discuss questions that have not been answered by the
project, but could be answered with future work?
• Detailed item (0 to 4). Did the team provide adequate reasons as to why these questions were left as
future work (e.g., they need more data, need more nuanced statistical techniques they haven’t learned,
etc.)? Are these future-work questions well-motivated given what the team has completed for this
project?

***
***

# Research Question 4: 

### Motivation
[PUT YOUR ANSWER HERE: 1 short paragraph connecting RQ1 to the real-world / scientific context.]

### Visualization 1
Detailed item (0 to 5). What is the quality of these graphs? Are these graphs easily readable,
interpretable, and properly labeled?
• Detailed item (0 to 5). Are the graphs well-motivated, given the research questions of the project?
In other words, do the graphs address/answer the research questions of the project, or do they only
provide tangential (or even irrelevant) details?
[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
 Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?
### Visualization 2

[[PUT YOUR ANSWER HERE: describe the variables and justify graph choice.]]

```{r}
#graph1
```

### Interpretation 
[PUT YOUR ANSWER HERE: additional interpretation, if applicable.]
Detailed item (0 to 4). What is the quality of these descriptions? After reading these descriptions, is it
crystal clear what is being displayed in the graph and what the main takeaways are?

### Conclusion

Does the report provide some clear conclusions that can be made from their
graphs and analyses?
• Detailed item (0 to 6). What is the accuracy of the claims being made? Are the claims well-supported
by the graphs and analyses presented? Are the conclusions well-aligned with the research questions of
the project?

### Further Research

Does the report discuss questions that have not been answered by the
project, but could be answered with future work?
• Detailed item (0 to 4). Did the team provide adequate reasons as to why these questions were left as
future work (e.g., they need more data, need more nuanced statistical techniques they haven’t learned,
etc.)? Are these future-work questions well-motivated given what the team has completed for this
project?

***
***

# Wrapping up





